The Error Estimate of Force Calculation in the Inhomogeneous Molecular Systems

### 2012-03-18 longRange.hetero/publish/main.tex
backup

In this paper, we consider the force error, bacause the applicaions
are based on the molecular dynamics simulation.  Suppose the system is
composed by $N$ charged particles located at $\v r_1, \v r_2, \cdots,
\v r_N$ with charges $q_1, q_2, \cdots, q_N$, respectively.  We study
the force error of an arbitrary charge $i$ by calculating the
difference between the exact force and calculated force exerting on
the charge $i$, which is called the \emph{error force}
\cite{wang2012} and denoted by $\Delta \v F(\v r_i)$.
% and is denoted by
% $\Delta \v F(\v r_i) = \v F(\v r_i) - \tilde{\v F}(\v r_i)$.
The maganitude of the force computation error 
% error of the force computation
is defined by the \emph{root mean square (RMS)} \emph{error}, which is
the square root of the second moment of the error force: $\mathcal
E(\v r_i) = \sqrt{\langle\vert\Delta\v F(\v r_i)\vert^2\rangle}$.  The
notation $\langle\cdot\rangle$ means ensemble average.  We also
estimate the the ensemble average of the error force (also called
\emph{mean error force}), namly $\langle\Delta\v F(\v r_i)\rangle$,
which plays an important role in the error analysis of the
inhomogeneous systems~\cite{wang2012}.

For error the pairwise interactions, we have the following theorem:

@@

  \begin{align}\label{eqn:tmp10}

@@

  \begin{align}\label{eqn:tmp11}

@@

    q_i^2\,[\v K\ast\rho_{q}]^2 (\v r_i) \\\label{eqn:tmp12}

@@

  \begin{align}\label{eqn:tmp13}

@@

    \bigg\rangle,\\\label{eqn:tmp14}

@@

    \bigg\rangle,\\\label{eqn:tmp15}
    C_{q^2}(\v r', \v r'')

@@

    \sum_{i\neq j}
    q_iq_j\delta(\v r'-\v r_i)\delta(\v r''-\v r_j)
    - \rho_q(\v r')\rho_q(\v r'').

@@

  $C_{q^2}(\v r', \v r'')$ are periodically extended to

@@

  \begin{align*} \nonumber

@@

@@

    [\,\rho_q(\v r')\rho_q(\v r'') + C_{q_2}(\v r',\v r'')\,]
    \,\d d\v r'\d d\v r''\\
    = \,&
    q_i^2\,[\v K\ast\rho_{q}]^2 (\v r_i) \\
    \d d\v r'\d d\v r''. \qquad\qed

@@

\item The function $\v K(\v r)$ is called \emph{the error force

@@

  Eqn.~\eqref{eqn:tmp13} and \eqref{eqn:tmp14} are called \emph{the
    first order charge density distribution (CDD)} and \emph{the
    second order charge density distribution},
  respectively. $C_{q^2}(\v r, \v r')$ defined in
  Eqn.~\eqref{eqn:tmp15} is the pairwise charge correlation function.

@@

  three terms on the r.h.s. of Eqn.~\eqref{eqn:tmp12} are called the

@@

  error, respectively. The homogeneity error and the inhomogeneity

@@

  However, the calculation of the correlation error costs $\mathcal
  O(N^2\log N)$, because it involves a twofold convolution. In this
  study, the contribution of the correlation error is neglicted.

@@

\item A non-zero mean error force~\eqref{eqn:tmp11} was proof to be
  harmful to the simulation in the case of the short-range force
  calculation~\cite{wang2012}.  In the case of the charge system, this
  term is non-zero only when the system is NOT locally neutral. For

@@

\subsection{Ewald summation and its error estimate}
The Ewald method divides the electrostatic interaction in to three
parts: the direct part, the reciprocal part and the correction
part. that is
\begin{equation}
E = E_{\textrm{dir}} + E_{\textrm{rec}} + E_{\textrm{corr}},
\end{equation}
where 
\begin {align}\label{eqn:tmp2}
E_{\textrm{dir}} & = \frac12 \sum^{\ast}_{\v n}
\sum_{i,j = 1}^{N} \frac{q_iq_j \textrm{erfc}(\beta \vert\v{r}_{ij} + \v{n}\vert)}
{\vert\v{r}_{ij} + \v{n}\vert} \\\label{eqn:tmp3}
E_{\textrm{rec}} & = \frac1{2\pi V} \sum_{\v m \neq 0}
\frac{\exp(-\pi^2\v m^2 / \beta^2)}{\v m^2} S(\v m) S(-\v m) \\\label{eqn:tmp4}
 E_{\textrm{corr}}& = -\frac\beta{\sqrt \pi} \sum_{i=1}^N q_i^2
\end {align}
The distance between the two particles is denoted by $\v r_{ij} = \v
r_i - \v r_j$, where $\v r_i$ and $\v r_j$ are positions of particle
$i$ and $j$, respectively.  The lattice in the real space is denoted
by $\v n = n_1 \v a_1 + n_2 \v a_2 + n_3 \v a_3$, where $\v a_\alpha,
\ \alpha=1,2,3$ are box vectors. the structure factor $S(\v m)$ is
defined by
\begin{equation}\label{sm1}
S(\v m) = \sum_{j=1}^N q_j \exp (2 \pi i \v m \cdot \v r_j),
\end{equation}
where $\v m = m_1 \v a_1^\ast + m_2 \v a_2^\ast + m_3 \v a_3^\ast$ is
the reciprocal space lattice vectors. $\v a_\alpha^\ast,\ \alpha = 1,
2, 3$ are conjugate reciprocal vectors of $\v a_\alpha$, defined by
$\v a_\alpha \cdot \v a_\beta^\ast = \delta_{\alpha\beta}$. and $V =
\v a_1 \cdot \v a_2 \times \v a_3$ is the volume of the box.

The complementary error function $\textrm{erfc}(r)$ in
Eqn.~\eqref{eqn:tmp2} decays exponentially as $r$ goes to infinity.
Therefore, it is a short-range interaction in the real space, which
can be calculated by the standard cut-off and neighbor list
method~\cite{frenkel02b} at a cost of $\mathcal O(N)$.  The reciprocal
part also decays exponentially as the maganitude of the Fourier mode
$\vert\v m\vert$ increases. Therefore, in practice, the infinit
summation in Eqn.~\eqref{eqn:tmp3} is truncated, and only a finite
summation is calculated:
% Both the direct and the reciprocal parts are calculated by cut-off in
% real and reciprocal space, respectively.
% For the reciprocal part the cut-offed energy reads
\begin{align}\label{eqn:tmp6}
  E^{\textrm{tr}}_{\textrm{rec}} & =
  % \frac1{2\pi V}
  \sum_{
    \begin{subarray}{c}
      \vert m_\alpha\vert < K_\alpha/2\\
      \v m\neq 0
    \end{subarray}}
  % \frac{\exp(-\pi^2\v m^2 / \beta^2)}{\v m^2} S(\v m) S(-\v m)   
  f(\v m)\, S(\v m) S(-\v m),
\end{align}
where, for short,
\begin{align}
  f(\v m) := \frac{\exp(-\pi^2 \v m^2 / \beta^2)}{2\pi V \v m^2},
\end{align}
and $K_\alpha$ is the number of Fourier modes used on direction
$\alpha$.  The truncated reciprocal force acting on particle $i$ is
\begin{align}\nonumber
  \v F^{\textrm{tr}}_{\textrm{rec}}(\v r_i)
  &= - 
  q_i 
  \sum_{
    \begin{subarray}{c}
      \vert m_\alpha\vert < K_\alpha/2\\
      \v m\neq 0
    \end{subarray}}
  \v g(\v m) \,
  S(-\v m)\,
  e^{2\pi i\v m\cdot\v r_i}, \\\label{eqn:tmp8}
  &= - 
  \sum_j   q_iq_j
  \sum_{
    \begin{subarray}{c}
      \vert m_\alpha\vert < K_\alpha/2\\
      \v m\neq 0
    \end{subarray}}
  \v g(\v m) \,
  e^{2\pi i\v m\cdot\v r_{ij}},
\end{align}
where
\begin{align}
  \v g(\v m) = -4 \pi \v m i f(\v m).
\end{align}

@@

Both the direct part and reciprocal part of the Ewald summation can be
written in the form of Eqn.~\eqref{eqn:tmp11}.  The direct
error force kernel is defined by:
% The direct part is a short-range interaction in real space, so it is
% usually treated by the cut-off method, which gives a error force
% kernel of
\begin{align}
  \v K^{\textrm{cut}}_{\textrm{dir}}(\v r) =
  \left\{
    \begin{aligned}
      &\,0, &\quad & r \leq r_c;\\
      &
      \Big[
      \frac{2\beta}{\sqrt\pi} e^{-\beta^2r^2} + \frac{\textrm{erfc}(\beta r)}{r}
      \Big]\frac{\v r}{r^2}
      ,& & r > r_c,
      % \frac{\textrm{erfc}(\beta\vert\v r\vert)}{\vert\v r\vert}
    \end{aligned}
  \right.
\end{align}
where $r = \vert\v r\vert$, and $r_c$ is the cut-off radius in the
real space.
% The reciprocal error force, namely the difference between
% the the exact and truncated reciprocal force, is
% \begin{align}
%   \Delta \v F^{\textrm{tr}}_{\textrm{rec}}(\v r_i)
%   = & - 
%   \sum_j\,q_iq_j
%   \sum_{\vert m_\alpha\vert \geq K_\alpha/2}
%   \v g(\v m) \,
%   e^{2\pi i\v m\cdot\v r_{ij}}.
% \end{align}
% Therefore,
The reciprocal error force kernel is given by:
\begin{align}
  \v K^{\textrm{tr}}_{\textrm{rec}}(\v r) =
  \sum_{
      \vert m_\alpha\vert \geq K_\alpha/2}
  \v g(\v m) \,
  e^{2\pi i\v m\cdot\v r}.
\end{align}
By the Therefore \ref{thm:tmp1}, the error estimate of the
truncated Ewald summation is straightforward.

@@

\subsection{The smooth particle mesh Ewald method}

The naive calculation of the reciprocal Ewald summation
\eqref{eqn:tmp6} will result in a computational cost of $\mathcal
O(N^2)$. It has been shown that with a careful tuning of parameter
$\beta$ and the reciprocal space truncation $K_\alpha$, the
computational cost is reduced to $\mathcal O(N^{3/2})$
\cite{perram1988asc}, which becomes not applicable when the system
size is larger than several hundreds of charged particles. Both the
particle mesh Ewald (PME) \cite{darden1993pme} and the smooth particle
mesh Ewald (SPME) \cite{essmann1995spm} methods are designed to reduce
the computational cost to $\mathcal O(N\log N)$ in nearly the same
way.  Since the SPME method is more precise \cite{deserno1998mue1} and
more popular than the PME, we will focus our discussion on the
former. All the error estimates develop by the current paper can be
easily extended to PME method.  The principle idea of the SPME method
is first to calculate the term $e^{2\pi i\v m\cdot\v r}$ on a uniform
lattice in the real space.  Therefore, all Fourier transforms can be
accerlerated by the fast Fourier transform (FFT), that is why the
computational cost of SPME is reduced to $\mathcal O(N\log N)$. Then,
for any particle position $\v r_i$, the value of $e^{2\pi i\v m\cdot\v
  r_i}$ is interpolated by the known values on the neighboring lattice
points.  The PME method uses Lagrangian interpolation, while the SPME
uses B-spline interpolation that is more precise, and has higher order
of smoothness.

>>>>

In this paper, we consider the force error, bacause the applicaions
are based on the molecular dynamics simulation.  Suppose the system is
composed by $N$ charged particles located at $\v r_1, \v r_2, \cdots,
\v r_N$ with charges $q_1, q_2, \cdots, q_N$, respectively.  We study
the force error of an arbitrary charge $i$ by calculating the
difference between the exact force and calculated force exerting on
the charge $i$, which is called the \emph{error force}
\cite{wang2012} and denoted by $\Delta \v F(\v r_i)$.
% and is denoted by
% $\Delta \v F(\v r_i) = \v F(\v r_i) - \tilde{\v F}(\v r_i)$.
The maganitude of the force computation error 
% error of the force computation
is defined by the \emph{root mean square (RMS)} \emph{error}, which is
the square root of the second moment of the error force: $\mathcal
E(\v r_i) = \sqrt{\langle\vert\Delta\v F(\v r_i)\vert^2\rangle}$.  The
notation $\langle\cdot\rangle$ means ensemble average.  We also
estimate the the ensemble average of the error force (also called
\emph{mean error force}), namly $\langle\Delta\v F(\v r_i)\rangle$,
which plays an important role in the error analysis of the
short-range inhomogeneous systems~\cite{wang2012}.

One of the main results of this research is the error estimate for
the error force that can be written as a summation of
pairwise interactions:

@@

  \begin{align}\label{eqn:thm-error-force}

@@

  \begin{align}\label{eqn:thm-meanf}

@@

    q_i^2\,[\v K\ast\rho_{q}]^2 (\v r_i) \\\label{eqn:thm-error}

@@

  \begin{align}\label{eqn:def-rhoq}

@@

    \bigg\rangle,\\\label{eqn:def-rhoq2}

@@

    \bigg\rangle,\\\label{eqn:def-rho2}
    \rho^{(2)}(\v r, \v r')

@@

    \sum_{j\neq k}
    q_jq_k\delta(\v r-\v r_j)\delta(\v r'-\v r_k)
    \bigg\rangle,\\\label{eqn:def-c}
    C(\v r, \v r')
    \rho^{(2)}(\v r, \v r')    
    - \rho_q(\v r)\rho_q(\v r').

@@

  $\rho^{(2)}(\v r, \v r')$ are periodically extended to

@@

  \begin{align} \nonumber

@@

    \\\label{eqn:prf-error-step3}

@@

    \rho^{(2)}(\v r', \v r'')
    \,\d d\v r'\d d\v r''\\\nonumber
    +
    q_i^2
    \int_{\mathbb R^3}
    \v K(\v r_i - \v r')\rho_{q}(\v r')\,\d d\v r'
    \bigg]^2
    \\ \nonumber
    C(\v r', \v r'')
    \,\d d\v r'\d d\v r''.\qquad\qed
    % = \,&
    % q_i^2\,[(\v K)^2\ast\rho_{q^2}] (\v r_i) + 
    % q_i^2\,[\v K\ast\rho_{q}]^2 (\v r_i) \\
    % &+
    % q_i^2\,\int_{\mathbb R^3\times\mathbb R^3}
    % \v K(\v r_i-\v r')\cdot
    % \v K(\v r_i-\v r'')\,
    % C_{q^2}(\v r', \v r'')\,
    % \d d\v r'\d d\v r''. \qquad\qed

@@

% \item
%   This theorem provides the error estimate for the interactions, whose
%   error force can be
%   written as a summation of pairwise interactions. 

@@

  Eqn.~\eqref{eqn:def-rhoq} and \eqref{eqn:def-rhoq2} are called \emph{the
    first order charge distribution} and \emph{the
    second order charge distribution},
  $\rho^{(2)}(\v r, \v r')$ in Eqn.~\eqref{eqn:def-rho2}
  is the \emph{pairwise charge distribution}.
  When the positions of any two different charges are independent,
  $\rho^{(2)}(\v r, \v r') = \frac{N-1}{N}\rho_q(\v r)\rho_q(\v r')$.
  Since $N$ is always a large number in real simulations, we take
  $\rho^{(2)}(\v r, \v r') = \rho_q(\v r)\rho_q(\v r')$ for convenience.
  $C_{q^2}(\v r, \v r')$ defined in
  Eqn.~\eqref{eqn:def-c} is the \emph{pairwise charge correlation function},
  which describes the strength of the correlation between  two
  different charges. When they are independent, $C_{q^2}(\v r, \v r')$
  vanishes, otherwise $C_{q^2}(\v r, \v r') \neq 0$.
  Most of the error estimates assumed the independency of the
  particles in the system~\cite{kolafa1992cutoff, deserno1998mue2, wang2010optimizing, wang2012},
  however, this could be problematic in the strongly correlated
  systems~\cite{deserno1998mue2, wang2010optimizing}.
  Also see Sec.~\ref{sec:example3} for example.
\item
  The \emph{pairwise form}
  of the interaction (see Eqn.~\eqref{eqn:thm-error-force}) is the key to derive
  the error estimates Eqn.~\eqref{eqn:thm-meanf} and \eqref{eqn:thm-error}.
  The function $\v K(\v r)$ is called \emph{the error force
  By setting $q_i = q_j = 1$, the theorem~\ref{thm:tmp1} directly proves
  the error estimates for the short-range interaction
  in inhomogeneous systems, which was derived in Ref.~\cite{wang2012}. 

@@

  three terms on the r.h.s. of Eqn.~\eqref{eqn:thm-error} are called the

@@

  error, respectively.
  This is denoted by:
  \begin{align}\label{eqn:error-split}
    \mathcal E^2_{\textrm{homo}}(\v r_i) +
    \mathcal E^2_{\textrm{inhomo}}(\v r_i) +
    \mathcal E_{\textrm{correlation}}(\v r_i).
  The homogeneity error and the inhomogeneity

@@

  However, the full calculation of the correlation error costs $\mathcal
  O(N^2\log N)$, because it involves a twofold convolution.
  % In this
  % study, the contribution of the correlation error is neglicted.

@@

\item A non-varnishing mean error force~\eqref{eqn:thm-meanf} was proved
  to be
  harmful to the calculation of the short-range interaction
  in inhomogeneity systems~\cite{wang2012}.
  In the case of the charged system, this
  term is non-zero only when the system is NOT locally neutral,
  i.e. a non-zero first order charge distribution. For

@@

\subsection{First neighbor approximation of the correlation
error}

In this section, we consider the first neighbor approximation of the
correlation error. Here ``first neighbors'' means the neighboring
atoms that are connected by the chemical bond.  In most molecular
force fields, such bonds are modeled by a rigid connection between two
atoms.  Therefore, the correlation between such an atom pair is very
strong. As an example, we consider a typical three point charge water
model TIP3P~\cite{jorgensen1983comparison}: each hydrogen atom carries
a charge of $q_{\textrm{H}} = +0.417e$ and each oxygen atom carries a charge of
$q_{\textrm{O}} = -0.834e$. Within one molecule, the H-O bond and H-O-H angle are
rigid. 
% In the system of water, the relative position of the oxygen and
% hydrogen atoms within one molecule is fixed. Therefore, according
% to this property, it is possible to improve the quality of the
% error estimate.
Starts from Eqn.~\eqref{eqn:prf-error-step3},
\begin{align} \nonumber
  \rho^{(2)}(\v r', \v r'')
  \,\d d\v r'\d d\v r'',
where the definition of $\rho^{(2)}(\v r', \v r'')$ is given by
Eqn.~\eqref{eqn:def-rho2}.
% \begin{align}
%   q_{q^2}(\v r', \v r'') =
%   \bigg\langle
%   \sum_{j\neq k}q_j q_k\delta(\v r' - \v r_j)\delta(\v r'' - \v r_k)
%   \bigg\rangle.
% \end{align}
We seperately consider the first neighbor contribution
and other contributions to
the pairwise density $\rho^{(2)}(\v r', \v r'')$.  Denoting
$\Omega_j = \{ \,k\,\vert\,
\textrm{atom $k$ is the first neighbor of atom $j$} \,\}$,
% the set of the first neighbors of the $j$-th atom by $\Omega_j$,
we have:
\begin{align}\label{eqn:split-rho-q2}
  \rho^{(2)}(\v r', \v r'')
  \sum_j\sum_{k\in\Omega_j}q_j q_k\delta(\v r' - \v r_j)\delta(\v r'' - \v r_k)
  +
  \sum_j\sum_{k\notin\Omega_j}q_j q_k\delta(\v r' - \v r_j)\delta(\v r'' - \v r_k)
  \bigg\rangle.
For any atom $k$ that is not the first neighbor of atom $j$, we
assume that its position is independent with atom $j$, then
second part on the r.h.s. of the Eqn.~\eqref{eqn:split-rho-q2} is
approximately $\frac{N-3}{N}\rho_q(\v r')\rho_q(\v r'')\approx
\rho_q(\v r')\rho_q(\v r'')$.
% If the number of atoms in the system is
% large, this should be a good approximation.
Therefore, the first part
on the r.h.s. of Eqn.~\eqref{eqn:split-rho-q2} is actually the
pairwise charge corrlation function:
\begin{align}\label{eqn:fnc-correlation-water}
  C_{q^2}(\v r', \v r'') =
  \sum_j\sum_{k\in\Omega_j}q_jq_k\delta(\v r' - \v r_j)\delta(\v r'' - \v r_k)
  \bigg\rangle.
% If connected by the chemical bond, the position of the first neighbors
% $k$ is not independent with the atom $j$, so their correlation is
% strong, and should be calculated explicitly to improve the quality of
% the error estimate.
By inserting Eqn.~\eqref{eqn:fnc-correlation-water} into
the correlation error, we have:
  \mathcal E_{\textrm{correlation}}(\v r_i)
  = q_i^2\sum_j\sum_{k\in\Omega_j}q_jq_k
  \v K(\v r_i - \v r_j)\cdot\v K(\v r_i - \v r_k)
  \big\rangle_{\v r_j,\v r_k}
Denoting the index set of the oxygen atoms by $\Omega_{\textrm{O}}$,
and the index set of the hydrogen atoms by $\Omega_{\textrm{H}}$, 
we have:
  \mathcal E_{\textrm{correlation}}(\v r_i)
  q_i^2\sum_{j\in\Omega_{\textrm{O}}}\sum_{k\in\Omega_j}q_jq_k
  \v K(\v r_i - \v r_j)\cdot\v K(\v r_i - \v r_k)
  \big\rangle_{\v r_j,\v r_k} \\ \nonumber
  q_i^2\sum_{j\in\Omega_{\textrm{H}}}\sum_{k\in\Omega_j}q_jq_k
  \v K(\v r_i - \v r_j)\cdot\v K(\v r_i - \v r_k)
  \big\rangle_{\v r_j,\v r_k} \\\nonumber
  q_i^2
  \int_{\mathbb R^3}
  \v K(\v r_i - \v r')\cdot
  q_{\textrm{H}}\v K(\v r_i - \v r' + \v s_{\textrm{H}_1})+
  q_{\textrm{H}}\v K(\v r_i - \v r' + \v s_{\textrm{H}_2})
  \big\rangle\,\rho_{\textrm{O}}(\v r')
  \d d\v r' \\\label{eqn:c-error-1}
  & \,+
  q_i^2
  \int_{\mathbb R^3}
  \v K(\v r_i - \v r')\cdot
  q_{\textrm{O}}\v K(\v r_i - \v r' + \v s_{\textrm{O}})+
  q_{\textrm{H}}\v K(\v r_i - \v r' + \v s_{\textrm{H}})
  \big\rangle\,\rho_{\textrm{H}}(\v r')
  \d d\v r'
If positions of the three atom within one molecule are
$\v r_{\textrm{H}_1}$, $\v r_{\textrm{H}_2}$ and $\v r_{\textrm{O}}$, then
$\v s_{\textrm{H}_1} = \v r_{\textrm{H}_1} - \v r_{\textrm{O}}$, 
$\v s_{\textrm{H}_2} = \v r_{\textrm{H}_2} - \v r_{\textrm{O}}$,
$\v s_{\textrm{O}} = \v r_{\textrm{O}} - \v r_{\textrm{H}_1}$ and
$\v s_{\textrm{H}} = \v r_{\textrm{H}_2} - \v r_{\textrm{H}_1}$.
The two ensemble averages in Eqn.~\eqref{eqn:c-error-1}
is taken over all possible water directions 
at position $\v r'$. As an approximation, we
assume these averages are  spacially uniform, namely independet with $\v r'$.
This assumption may be  problematic
in the regions where the preference of water  drection is different
with others.
% In the example considered by the present paper in section~\ref{sec:tmp3},
% it is shown to be a good approximation to obtain the estimate of the
% correlation error.
One possible way to calculate the ensemble averages in
Eqn.~\eqref{eqn:c-error-1} is by the Fourier transform
of the error force kernel:
\begin{align} \nonumber
  &\big\langle
  q_{\textrm{H}}\v K(\v r_i - \v r' + \v s_{\textrm{H}_1})+
  q_{\textrm{H}}\v K(\v r_i - \v r' + \v s_{\textrm{H}_2})
  \big\rangle \\ \nonumber
  \frac1V\bigg\langle
  \sum_{\v m}
  q_{\textrm{H}}\hat{\v K}(\v m)
  e^{2\pi i\v m\cdot(\v r_i - \v r' + \v s_{\textrm{H}_1})} +
  \sum_{\v m}
  q_{\textrm{H}}\hat{\v K}(\v m)
  e^{2\pi i\v m\cdot(\v r_i - \v r' + \v s_{\textrm{H}_2})}
  \bigg\rangle \\\nonumber
  \frac1V\sum_{\v m}
  \hat{\v K}(\v m)
  e^{2\pi i\v m\cdot(\v r_i - \v r')}
  \,  
  \langle
  q_{\textrm{H}}\,e^{2\pi i\v m\cdot\v s_{\textrm{H}_1}} +
  q_{\textrm{H}}\,e^{2\pi i\v m\cdot\v s_{\textrm{H}_2}}
  \rangle,
where ``$\wedge$'' denotes the forward Fourier transform. By denoting
  \hat T_{\textrm{O}}(\v m)
  \langle
  q_{\textrm{H}}\,e^{2\pi i\v m\cdot\v s_{\textrm{H}_1}} +
  q_{\textrm{H}}\,e^{2\pi i\v m\cdot\v s_{\textrm{H}_2}}
  \rangle,\\
  \hat T_{\textrm{H}}(\v m)
  \langle
  q_{\textrm{O}}\,e^{2\pi i\v m\cdot\v s_{\textrm{O}}} +
  q_{\textrm{H}}\,e^{2\pi i\v m\cdot\v s_{\textrm{H}}}
  \rangle.
The Eqn.~\eqref{eqn:c-error-1} becomes:
  \mathcal E_{\textrm{correlation}}(\v r_i)
  =&\,
  q_i^2
  \int_{\mathbb R^3}
  \v K(\v r_i - \v r')\cdot
  (\hat T_{\textrm{O}}\hat{\v K})^{\vee}(\v r_i - \v r')
  \,\rho_{\textrm{O}}(\v r')
  \,\d d\v r' \\\label{eqn:c-error-water}
  & \,+
  q_i^2
  \int_{\mathbb R^3}
  \v K(\v r_i - \v r')\cdot
  (\hat T_{\textrm{H}}\hat{\v K})^{\vee}(\v r_i - \v r')
  \,\rho_{\textrm{H}}(\v r')
  \,\d d\v r',
where ``$\vee$'' denotes the backward Fourier transform.
In the Fourier space, the error force kernel $\v K(\v r)$ is
influenced by the prefactors $T_{\textrm{O}}$ and $T_{\textrm{H}}$,
which are accounting for the atomic correlation within one water
molecule.

% For the second term, do variable change $\v r'' = \v r' + \v s$, we have:
% \begin{align}
%   \textrm{II} = & 
%   q_i^2
%   \int_{\mathbb R^3}\d d\v r'\v K(\v r_i - \v r')
%   \int_{\mathbb R^3}\d d\v s\,
%   \v K(\v r_i - \v r' - \v s)\,
%   \rho_{q^2}(\v r', \v r' + \v s).
% \end{align}
% Split the integral region of the variable $\v s$ into the first
% neighbor shell $\mathcal B(\v r', R_s)$ and the reset $\mathbb R -
% \mathcal B(\v r', R_s)$. In the case of water, the first neighbor
% shell means the region where a single molecule occupies. Within this
% shell, the relative position of the oxygen and two hydrogens are
% always fixed. Outside of the first neighbor shell, the position of
% oxygens and hydrogens are belongs to different molecules, so the their
% positions are assumed to be independent. In most cases, the first
% neighbor shell is small comparing to the simulation box size,
% therefore,
% \begin{align}\nonumber
%   \textrm{II}_{\textrm{out}} = & \,
%   q_i^2
%   \int_{\mathbb R^3}\d d\v r'\v K(\v r_i - \v r')
%   \int_{\mathbb R^3 - \mathcal B(\v r', s)}\d d\v s\,
%   \v K(\v r_i - \v r' - \v s)\,
%   \rho_{q^2}(\v r', \v r' + \v s) \\ \nonumber
%   = &\,
%   q_i^2
%   \int_{\mathbb R^3}\d d\v r'\v K(\v r_i - \v r') \rho_q(\v r')
%   \int_{\mathbb R^3 - \mathcal B(\v r', s)}\d d\v s\,
%   \v K(\v r_i - \v r' - \v s)\,
%   \rho_q(\v r' + \v s) \\
%   \approx &\,
%   q_i^2
%   [\v K\ast\rho_{q}]^2 (\v r_i)
% \end{align}

\subsection{Ewald summation and its error estimate}
The Ewald method divides the electrostatic interaction in to three
parts: the direct part, the reciprocal part and the correction
part:
\begin{equation}
E = E_{\textrm{dir}} + E_{\textrm{rec}} + E_{\textrm{corr}},
\end{equation}
where 
\begin {align}\label{eqn:tmp2}
E_{\textrm{dir}} & = \frac12 \sum^{\ast}_{\v n}
\sum_{i,j = 1}^{N} \frac{q_iq_j \textrm{erfc}(\beta \vert\v{r}_{ij} + \v{n}\vert)}
{\vert\v{r}_{ij} + \v{n}\vert} \\\label{eqn:tmp3}
E_{\textrm{rec}} & = \frac1{2\pi V} \sum_{\v m \neq 0}
\frac{\exp(-\pi^2\v m^2 / \beta^2)}{\v m^2} S(\v m) S(-\v m) \\\label{eqn:tmp4}
 E_{\textrm{corr}}& = -\frac\beta{\sqrt \pi} \sum_{i=1}^N q_i^2
\end {align}
The distance between the two particles is denoted by $\v r_{ij} = \v
r_i - \v r_j$.  The lattice in the real space is denoted
by $\v n = n_1 \v a_1 + n_2 \v a_2 + n_3 \v a_3$, where $\v a_\alpha,
\ \alpha=1,2,3$ are box vectors. the structure factor $S(\v m)$ is
defined by
\begin{equation}\label{sm1}
S(\v m) = \sum_{j=1}^N q_j \exp (2 \pi i \v m \cdot \v r_j),
\end{equation}
where $\v m = m_1 \v a_1^\ast + m_2 \v a_2^\ast + m_3 \v a_3^\ast$ is
the reciprocal space lattice vectors. $\v a_\alpha^\ast,\ \alpha = 1,
2, 3$ are conjugate reciprocal vectors of $\v a_\alpha$, defined by
$\v a_\alpha \cdot \v a_\beta^\ast = \delta_{\alpha\beta}$. and $V =
\v a_1 \cdot \v a_2 \times \v a_3$ is the volume of the box.

The complementary error function $\textrm{erfc}(r)$ in
Eqn.~\eqref{eqn:tmp2} decays exponentially as $r$ goes to infinity.
Therefore, it is a short-range interaction in the real space, which
can be calculated by the standard cut-off and neighbor list
method~\cite{frenkel02b} at a cost of $\mathcal O(N)$.  The reciprocal
part also decays exponentially as the maganitude of the Fourier mode
$\vert\v m\vert$ increases. Therefore, in practice, the infinit
summation in Eqn.~\eqref{eqn:tmp3} is truncated, and only a finite
summation is calculated:
% Both the direct and the reciprocal parts are calculated by cut-off in
% real and reciprocal space, respectively.
% For the reciprocal part the cut-offed energy reads
\begin{align}\label{eqn:tmp6}
  E^{\textrm{tr}}_{\textrm{rec}} & =
  % \frac1{2\pi V}
  \sum_{
    \begin{subarray}{c}
      \vert m_\alpha\vert < K_\alpha/2\\
      \v m\neq 0
    \end{subarray}}
  % \frac{\exp(-\pi^2\v m^2 / \beta^2)}{\v m^2} S(\v m) S(-\v m)   
  f(\v m)\, S(\v m) S(-\v m),
\end{align}
where, for short,
\begin{align}
  f(\v m) := \frac{\exp(-\pi^2 \v m^2 / \beta^2)}{2\pi V \v m^2},
\end{align}
and $K_\alpha$ is the number of Fourier modes used on direction
$\alpha$.  The truncated reciprocal force acting on particle $i$ is
\begin{align}\nonumber
  \v F^{\textrm{tr}}_{\textrm{rec}}(\v r_i)
  &= 
  q_i 
  \sum_{
    \begin{subarray}{c}
      \vert m_\alpha\vert < K_\alpha/2\\
      \v m\neq 0
    \end{subarray}}
  \v g(\v m) \,
  S(-\v m)\,
  e^{2\pi i\v m\cdot\v r_i}, \\\label{eqn:tmp8}
  &= 
  \sum_j   q_iq_j
  \sum_{
    \begin{subarray}{c}
      \vert m_\alpha\vert < K_\alpha/2\\
      \v m\neq 0
    \end{subarray}}
  \v g(\v m) \,
  e^{2\pi i\v m\cdot\v r_{ij}},
\end{align}
where
\begin{align}
  \v g(\v m) = -4 \pi \v m i f(\v m).
\end{align}

@@

Both the direct part and reciprocal part of the Ewald summation can be
written in the pairwise form of Eqn.~\eqref{eqn:thm-error-force}.  The direct
error force kernel is defined by:
% The direct part is a short-range interaction in real space, so it is
% usually treated by the cut-off method, which gives a error force
% kernel of
\begin{align}
  \v K^{\textrm{cut}}_{\textrm{dir}}(\v r) =
  \left\{
    \begin{aligned}
      &\,0, &\quad & r \leq r_c;\\
      &
      \Big[
      \frac{2\beta}{\sqrt\pi} e^{-\beta^2r^2} + \frac{\textrm{erfc}(\beta r)}{r}
      \Big]\frac{\v r}{r^2}
      ,& & r > r_c,
      % \frac{\textrm{erfc}(\beta\vert\v r\vert)}{\vert\v r\vert}
    \end{aligned}
  \right.
\end{align}
where $r = \vert\v r\vert$, and $r_c$ is the cut-off radius in the
real space.
% The reciprocal error force, namely the difference between
% the the exact and truncated reciprocal force, is
% \begin{align}
%   \Delta \v F^{\textrm{tr}}_{\textrm{rec}}(\v r_i)
%   = & - 
%   \sum_j\,q_iq_j
%   \sum_{\vert m_\alpha\vert \geq K_\alpha/2}
%   \v g(\v m) \,
%   e^{2\pi i\v m\cdot\v r_{ij}}.
% \end{align}
% Therefore,
The reciprocal error force kernel is given by:
\begin{align}
  \v K^{\textrm{tr}}_{\textrm{rec}}(\v r) =
  \sum_{
      \vert m_\alpha\vert \geq K_\alpha/2}
  \v g(\v m) \,
  e^{2\pi i\v m\cdot\v r}.
\end{align}
By the Therefore \ref{thm:tmp1}, the error estimate of the
truncated Ewald summation is straightforward.

@@

\subsection{The smooth particle mesh Ewald method \&
the staggered mesh Ewald method}

The naive calculation of the reciprocal Ewald summation
\eqref{eqn:tmp6} will result in a computational cost of $\mathcal
O(N^2)$. It has been shown that with a careful tuning of parameter
$\beta$, real space cut-off radius $r_c$
and the reciprocal space truncation $K_\alpha$, the
optimal computational cost is $\mathcal O(N^{3/2})$~\cite{perram1988asc},
which becomes not applicable when the system
size is larger than several hundreds of charged particles. Both the
particle mesh Ewald (PME) \cite{darden1993pme} and the smooth particle
mesh Ewald (SPME) \cite{essmann1995spm} methods are designed to reduce
the computational cost to $\mathcal O(N\log N)$ in nearly the same
way.  Since the SPME method is more precise \cite{deserno1998mue1} and
more popular than the PME, we will focus our discussion on the
former. All the error estimates develop by the current paper can be
easily extended to the PME method.  The principle idea of the SPME method
is first to calculate the term $e^{2\pi i\v m\cdot\v r}$ on a uniform
lattice in the real space.  Therefore, all Fourier transforms can be
accerlerated by the fast Fourier transform (FFT) algoritm, that is why the
computational cost of SPME is reduced to $\mathcal O(N\log N)$. Then,
for any particle position $\v r_i$, the value of $e^{2\pi i\v m\cdot\v
  r_i}$ is interpolated by the known values on the neighboring lattice
points.  The PME method uses Lagrangian interpolation, while the SPME
uses B-spline interpolation,  which is more precise, and has higher order
of smoothness.
